\textbf{Microbiota predict \emph{Clostridium difficile} severity in germ-free mice colonized with human feces}

### 2022-04-25 submission/manuscript.Rmd

***C. difficile* is able to infect germ-free mice colonized with human fecal microbial communities without antibiotics.**  To produce gut microbiomes with greater variation than those found in conventional mouse colonies, we colonized germ-free mice with bacteria from human feces [@Nagpal2018]. We inoculated germ-free C57BL/6 mice with homogenized feces from each of 15 human fecal samples via oral gavage. These human fecal samples were selected because they represented diverse community structures based on community clustering [@Schubert2014]. The gut communities were allowed to colonize for two weeks post-inoculation [@Gillilland2012]. We then surveyed the bacterial members of the gut communities by 16S rRNA gene sequencing of murine fecal pellets (Figure 1A). The bacterial communities from each mouse grouped more closely to those communities from mice that received the same human fecal donor community than to the mice who received a different human fecal donor community (Figure 1B). The communities were primarily composed of populations of *Clostridia*, *Bacteroidia*, *Erysipelotrichia*, *Bacilli*, and *Gammaproteobacteria*. However, the gut bacterial communities of each donor group of mice harbored unique relative abundance distributions of the shared bacterial classes.

@@

**Statistical analysis and modeling.** To compare community structures, we calculated Yue and Clayton dissimilarity matrices ($\theta$~YC~) in mothur [@Yue2005]. For this calculation, we averaged of 1000 sub-samples, or rarified, samples to 2,107 sequence reads per sample to limit uneven sampling biases. We tested for differences in individual taxonomic groups that would explain the outcome differences with LEfSe [@Segata2011] in mothur (default parameters, LDA > 4). We tested for differences in temporal trends through fitting a linear model to each OTU and testing for differences between histopathological summary scores with LEfSe [@Segata2011] in mothur (default parameters, LDA > 3). Remaining statistical analysis and data visualization was performed in R (v4.0.5) with the tidyverse package (v1.3.1). We tested for significant differences in $\beta$-diversity ($\theta$~YC~), histopathological scores, and toxin activity using the Wilcoxon rank sum test, non-unimodality to non-moribund histopathological summary score using Hartigans' dip test, and  toxin detection in mice using the Pearson's Chi-square test. We used Spearman's correlation to identify which OTUs that had a correlation between their relative abundance and the histopathologic summary score. *P* values were then corrected for multiple comparisons with a Benjamini and Hochberg adjustment for a type I error rate of 0.05 [@Benjamini1995]. We built L2 logistic regression models using the mikropml package [@Topcuoglu2021]. Sequence counts were summed by taxonomic ranks from day 0 samples, normalized by centering to the feature mean and scaling by the standard deviation, and features positively or negatively correlated were collapsed into a single feature. We ran our models using alpha = 0 and lambda values of 1e-0, 1e1, 1e2, 2e2, 3e2, 4e2, 5e2, 6e2, 7e2, 8e2, 9e2, 1e3, 1e4 with a split of 80% of the data for training and 20% of the data for testing. Lastly, we did not compare murine communities to donor community or clinical data because germ-free mice colonized with non-murine fecal communities have been shown to more closely resemble the murine communities than the donor species community [@Rawls2006]. Furthermore, it is not our intention to make any inferences regarding human associated bacteria and their relationship with human CDI outcome.

@@

Conceptualization: N.A.L., A.M.S., K.J.F., P.D.S.; Data curation: N.A.L., K.J.F.; Formal analysis: N.A.L., K.J.F., J.L.L., I.L.B.; Investigation: N.A.L., A.M.S., H.S., I.L.B., V.B.Y., P.D.S.; Methodology: N.A.L., A.M.S., K.J.F., J.L.L., H.S., I.L.B., V.B.Y., P.D.S.; Resources: N.A.L., A.M.S., P.D.S.; Software: NAL; Visualization: N.A.L., K.J.F., P.D.S.; Writing - original draft: N.A.L.; Writing - review & editing: N.A.L., A.M.S., K.J.F., J.L.L., H.S., I.L.B., V.B.Y., P.D.S.; Funding acquisition: V.B.Y.; Project administration: P.D.S.; Supervision: P.D.S.

>>>>

***C. difficile* is able to infect germ-free mice colonized with human fecal microbial communities without antibiotics.**  To produce gut microbiomes with greater variation than those found in conventional mouse colonies, we colonized germ-free mice with bacteria from human feces [@Nagpal2018]. We inoculated germ-free C57BL/6 mice with homogenized feces from each of 15 human fecal samples via oral gavage. These human fecal samples were selected because they represented diverse community structures based on community clustering [@Schubert2014]. After the gut communities had colonized for two weeks, we confirmed them to be *C. difficile* negative by culture [@Gillilland2012]. We then surveyed the bacterial members of the gut communities by 16S rRNA gene sequencing of murine fecal pellets (Figure 1A). The bacterial communities from each mouse grouped more closely to those communities from mice that received the same human fecal donor community than to the mice who received a different human fecal donor community (Figure 1B). The communities were primarily composed of populations of *Clostridia*, *Bacteroidia*, *Erysipelotrichia*, *Bacilli*, and *Gammaproteobacteria*. However, the gut bacterial communities of each donor group of mice harbored unique relative abundance distributions of the shared bacterial classes.

@@

**Statistical analysis and modeling.** To compare community structures, we calculated Yue and Clayton dissimilarity matrices ($\theta$~YC~) in mothur [@Yue2005]. For this calculation, we averaged of 1000 sub-samples, or rarified, samples to 2,107 sequence reads per sample to limit uneven sampling biases. We tested for differences in individual taxonomic groups that would explain the outcome differences with LEfSe [@Segata2011] in mothur (default parameters, LDA > 4). We tested for differences in temporal trends through fitting a linear model to each OTU and testing for differences between histopathological summary scores with LEfSe [@Segata2011] in mothur (default parameters, LDA > 3). Remaining statistical analysis and data visualization was performed in R (v4.0.5) with the tidyverse package (v1.3.1). We tested for significant differences in $\beta$-diversity ($\theta$~YC~), histopathological scores, and toxin activity using the Wilcoxon rank sum test, non-unimodality to non-moribund histopathological summary score using Hartigans' dip test, and  toxin detection in mice using the Pearson's Chi-square test. We used Spearman's correlation to identify which OTUs that had a correlation between their relative abundance and the histopathologic summary score. *P* values were then corrected for multiple comparisons with a Benjamini and Hochberg adjustment for a type I error rate of 0.05 [@Benjamini1995]. We built L2 logistic regression models using the mikropml package [@Topcuoglu2021]. Sequence counts were summed by taxonomic ranks from day 0 samples, normalized by centering to the feature mean and scaling by the standard deviation, and features positively or negatively correlated were collapsed into a single feature. For each L2 logistic regression model, we ran 100 random iterations using values of 1e-0, 1e1, 1e2, 2e2, 3e2, 4e2, 5e2, 6e2, 7e2, 8e2, 9e2, 1e3, 1e4 for the L2 regularization penalty with a split of 80% of the data for training and 20% of the data for testing. Lastly, we did not compare murine communities to donor community or clinical data because germ-free mice colonized with non-murine fecal communities have been shown to more closely resemble the murine communities than the donor species community [@Rawls2006]. Furthermore, it is not our intention to make any inferences regarding human associated bacteria and their relationship with human CDI outcome.

@@

Conceptualization: N.A.L., A.M.S., K.J.F., P.D.S.; Data curation: N.A.L., K.J.F.; Formal analysis: N.A.L., K.J.F., J.L.L., I.L.B.; Investigation: N.A.L., A.M.S., H.S., I.L.B., V.B.Y., P.D.S.; Methodology: N.A.L., A.M.S., K.J.F., J.L.L., H.S., I.L.B., V.B.Y., P.D.S.; Resources: N.A.L., A.M.S., P.D.S.; Software: NAL; Visualization: N.A.L., K.J.F., P.D.S.; Writing - original draft: N.A.L.; Writing - review & editing: N.A.L., A.M.S., K.J.F., J.L.L., H.S., I.L.B., V.B.Y., P.D.S.; Funding acquisition: V.B.Y.; Project administration: P.D.S.; Supervision: P.D.S.
\newpage
